---
title: <center><h1>2018R1 High-Dimensional Data Analysis (STAT5103) Assignment 1</h1></center><br />
author: <center>Yiu Chung WONG 1155017920</center>
output:
  html_document:
    keep_md: yes
  word_document: default
  pdf_document: default
--- 
<br />
<br />

###1. Let
$$\mathbf{A} = \left(\begin{array}
{rrr}
1 & 2 & 3 \\
2 & 1 & 2 \\
3 & 2 & 1
\end{array}\right)$$


```{r, echo = F, results = 'hide', message=FALSE, warning=FALSE}
require("knitr")
library(dplyr)
opts_chunk$set(echo = TRUE, cache=TRUE)
options(width = 90)
```

```{r}
A <- matrix(c(1,2,3,2,1,2,3,2,1), 3, 3)
A
```
<br />

####(a) Find the eigenvalues ($\lambda_{1}$, $\lambda_{2}$, $\lambda_{3}$) and their corresponding normalized eigenvectors ($\mathrm{x_1}$, $\mathrm{x_2}$, $\mathrm{x_3}$)  
```{r}
eigen(A)
```

<br />

####(b) Let **P** = ($x_{1}$, $x_{2}$, $x_{3}$), a matrix formed by using the eigenvectors as columns. Find **PDP¡¬** where **D** is a diagonal matrix with the eigenvalues ($\lambda_{1}$, $\lambda_{2}$, $\lambda_{3}$) on its diagonal. 
```{r}
P <- eigen(A)$vectors
D <- diag(eigen(A)$values)
P %*% D %*% solve(P)
```

<br />

####(c) Find the singular value decomposition of **A**. 
```{r}
svd(A)
```

<br />

###2. Let the eigenvalue-eigenvector pairs of a 2x2 symmetric matrix **A** be 
<br />

\[
  \left(3,\;\left(\begin{array}{cc} 
1^{}/{\sqrt{2}}\\
-1^{}/{\sqrt{2}}
\end{array}\right) \right)  
\;and\;
  \left(5,\;\left(\begin{array}{cc} 
1^{}/{\sqrt{2}}\\
1^{}/{\sqrt{2}}
\end{array}\right) \right)
\]
<br />


####(a) Find **A**. 
```{r}
x = 1/sqrt(2)
P = matrix(c(x, -x, x, x), 2, 2)
D = diag(c(3,5))
A = P %*% D %*% solve(P)
A
```

<br />

####(b) Find the determinant of **A**. 
```{r}
det(A)
```

<br />

####(c) Find the trace of **A**. 
```{r}
sum(diag(A))
```

<br />

###3. Using $x_{3}$ - $x_{9}$ of the US Crime Data, compute 
<br />

####(a) the sample mean ($\bar{\mathrm{x}}$),
```{r}
crime <- readxl::read_excel("uscrime.xlsx", sheet = 1, skip = 1)
sapply(crime[, 3:9], mean)
```
<br />

####(b) the sample covariance matrix (**S**), and
<br />
<center>$cov_{x,y}=\frac{\sum_{i=1}^{N}(x_{i}-\bar{x})(y_{i}-\bar{y})}{N-1}$</center>
<br />

```{r}
cov(crime[, 3:9]) %>% round(., 2)
```
<br />

####(c) the sample correlation matrix (**R**).
<br />
<center>$r =\frac{\sum ^n _{i=1}(x_i - \bar{x})(y_i - \bar{y})}{\sqrt{\sum ^n _{i=1}(x_i - \bar{x})^2} \sqrt{\sum ^n _{i=1}(y_i - \bar{y})^2}}$</center>
<br />

```{r}
cor(crime[, 3:9]) %>% round(., 2)
```