---
title: <center><h1>2019R2 STAT6108 Assignment 5</h1></center><br />
author: <center>Yiu Chung WONG 1155017920</center>
output:
  pdf_document: default
  html_document:
    keep_md: yes
  word_document: default
--- 
<br />
<br />

```{r, echo = F, results = 'hide', message=FALSE, warning=FALSE}
library("knitr")
opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
options(width = 100)
library("lavaan")
library("semTools")
library("dplyr")

knitr::knit_hooks$set(inline = function(x) {
  x <- sprintf("%1.3f", x)
  paste(x, collapse = ", ")
})
```

```{r}
# import data
male_cov <- getCov(scan('hw5_male(2020).cov'))
female_cov <- getCov(scan('hw5_female(2020).cov'))
```

### Model Setup

```{r}
n_male <- 204
n_female <- 201

sample.cov <- list(male = male_cov, female = female_cov)
sample.nobs <-  list(n_male, n_female)

model <- "
# measurment model
Motivation =~ V1 + V2;
JobCommitment =~ V3 + V4;
JobSatisfaction =~ V5 + V6 + V7;

# structural model
JobCommitment ~  Motivation;
JobSatisfaction ~ JobCommitment +  Motivation;
"
```
<br />
\newpage 

## a) 

###  multiple group model fitting

```{r, results='hide'}
base_syntax <- measEq.syntax(configural.model = model,
                             sample.cov = sample.cov,
                             sample.nobs = sample.nobs,
                             auto.var = TRUE,
                             ID.fac = "auto.fix.first")

# user-defined parameters
indirect_male <- "indirect_male := beta.2_1.g1 * beta.3_2.g1"
indirect_female <- "indirect_female := beta.2_1.g2 * beta.3_2.g2"

# add user-defined parameters to the @constraints slot
base_syntax@constraints <- c(base_syntax@constraints, indirect_male, indirect_female)

# train model
base_model <- lavaan(model = as.character(base_syntax),
                     sample.cov = sample.cov,
                     sample.nobs = sample.nobs)

# goodness-of-fit measures and parameter estimates
fits_base <- fitMeasures(base_model,
                        fit.measures = c("chisq", "pvalue", "tli", "cfi", "rmsea", "srmr"))
para_base <- parameterEstimates(base_model)
```

* $H_0$: $\Sigma = \Sigma(\theta)$ 
* $\chi^2$ test statistics: `r fits_base["chisq"]`; p-value: `r fits_base["pvalue"]`.
* $H_0$ is rejected at $\alpha = .05$.

* Male test statistic: `r lavInspect(base_model, "test")$standard$stat.group[1]`
* Female test statistic: `r lavInspect(base_model, "test")$standard$stat.group[2]`

Female test statistic slightly bigger. This indicates the proposed structure does not fit the female sample as well on as the male sample.

**We evaluate the indicies according to Hooper, Coughlan, & Mullen (2008).**

* NNFI: `r fits_base["tli"]` > $0.95$
* CFI: `r fits_base["cfi"]` > $0.95$
* RMSEA: `r fits_base["rmsea"]` < $0.07$
* SRMR: `r fits_base["srmr"]` < $0.08$

Goodness-of-fit indices pass threashold across all four major measures

## b)

* Wald test p-value for male direct effect $B_{31}^{(F)}$: `r filter(para_base, label == "beta.3_1.g1") %>% select(.,  pvalue)`
* Wald test p-value for female direct effect $B_{31}^{(M)}$: `r filter(para_base, label == "beta.3_1.g2") %>% select(.,  pvalue)`
* Neither are sigificant at $\alpha = .05$

## c) 

* Wald test p-value for male indirect effect $B_{21}^{(M)} {\cdot} B_{32}^{(M)}$: `r filter(para_base, label == "indirect_male") %>% select(.,  pvalue)`
* Wald test p-value for female indirect effect $B_{21}^{(F)} {\cdot} B_{32}^{(F)}$: `r filter(para_base, label == "indirect_female") %>% select(.,  pvalue)`
* Both are sigificant at $\alpha = .05$.

For both groups, since only indirect effects are statistically sigificant, the proposed structure is fully mediated.
\newpage 

## d)
```{r}
invariance_syntax <- measEq.syntax(configural.model = base_model,
                                   sample.cov = sample.cov,
                                   sample.nobs = sample.nobs,
                                   ID.fac = "auto.fix.first",
                                   group.equal = "loadings") # fix factor loadings

# user-defined parameters
variance_diff <- "variance_diff := psi.1_1.g1 - psi.1_1.g2"
direct_diff <- "direct_diff := beta.3_1.g1 - beta.3_1.g2"
indirect_diff <- "indirect_diff := indirect_male - indirect_female"

# add user-defined parameters to the @constraints slot
invariance_syntax@constraints <- c(base_syntax@constraints,
                                   variance_diff,
                                   direct_diff,
                                   indirect_diff)

# train model
invariance_model <- lavaan(model = as.character(invariance_syntax),
                     sample.cov = sample.cov,
                     sample.nobs = sample.nobs)

# goodness-of-fit measures and parameter estimates
fits_invar <- fitMeasures(invariance_model,
                        fit.measures = c("chisq", "pvalue", "tli", "cfi", "rmsea", "srmr"))
para_invar <- parameterEstimates(invariance_model)
```

#### Liklihood ratio test
```{r}
li_test <- compareFit(base_model, invariance_model)
li_test
```
```{r, echo=FALSE, results='hide'}
li_test <- as.data.frame(li_test@nested)
```



**The $\chi^2$ value is the likelihood-ratio test statistic**

* $H_0$: $\Lambda^{(M)} = \Lambda^{(F)}$ 
* p-value for likelihood ratio test is: `r li_test[["Pr(>Chisq)"]][which(row.names(li_test) == "invariance_model")]`
* At $\alpha = .05$, $H_0$ not rejected; weak invariance holds.

**Goodness-of-fit of weak invariance model**

* $H_0$: $\Sigma = \Sigma(\theta)$ 
* $\chi^2$ test statistics: `r fits_invar["chisq"]`; p-value: `r fits_invar["pvalue"]`. $H_0$ is not rejected at $\alpha = .05$.
* NNFI: `r fits_invar["tli"]` > $0.95$
* CFI: `r fits_invar["cfi"]` > $0.95$
* RMSEA: `r fits_invar["rmsea"]` < $0.07$
* SRMR: `r fits_invar["srmr"]` < $0.08$

The weak invariance model has goodness-of-fit indices pass threashold across all four major measures. $\chi^2$ test of differences is not sigificant. 
The proposed structure fits the data well and better than the configural model.

## e)
* $H_0$: $\psi_{11}^{(F)} \leq \psi_{11}^{(M)}$
* Point estimate of male variance is greater than that of female by: `r filter(para_invar, label == "variance_diff") %>% select(.,  est)`
* One sided Wald test p-value for point estimate of difference in motivation variance: `r 0.5 * (filter(para_invar, label == "variance_diff") %>% select(.,  pvalue) )`
* At $\alpha = .05$, $H_0$ rejected.
* There is sigificant evidence indicating male employees are more heterogeneous in terms of their motivation than their female counterparts.

## f)
* $H_0$: $B_{31}^{(M)} = B_{31}^{(F)}$
* Wald test p-value for point estimate of difference in direct effect: `r filter(para_invar, label == "direct_diff") %>% select(.,  pvalue)`
* At $\alpha = .05$, $H_0$ not rejected. 
* There is no sigificant evidence indicating direct effect of motivation on job satisfaction is not the same in the two groups.

## g)
* $H_0$: $B_{21}^{(M)} {\cdot} B_{32}^{(M)} = B_{21}^{(F)} {\cdot} B_{32}^{(F)}$ 
* Wald test p-value for point estimate of difference in direct: `r filter(para_invar, label == "indirect_diff") %>% select(.,  pvalue)`
* At $\alpha = .05$, $H_0$ not rejected. 
* There is no sigificant evidence indicating mediation effect is moderated by gender.



