---
title: <center><h1> 2019R1 Discrete Data Analysis (STAT5107) Assignment 2</h1></center><br />
author: <center>Yiu Chung WONG 1155017920</center>
output:
  pdf_document: default
  html_document:
    keep_md: yes
  word_document: default
--- 
<br />
<br />

```{r, echo = FALSE,results = 'hide'}
gc()
rm(list = ls())
```


```{r, echo = F, results = 'hide', message=FALSE, warning=FALSE}
library("knitr")
opts_chunk$set(echo = TRUE, cache=TRUE, message=FALSE, warning = FALSE)
options(width = 100)
library("ggplot2")
library("GGally")
library("dplyr")
```

```{r}
set.seed(5107);
```

#### 1a.

* relative risk
<br />

#### 1bi.
```{r}
(1-.45) / 1
```
<br />

#### 1bii.
```{r}
1 / (1-.45)
```
<br />

#### 2.

$$
\begin{aligned}
 P(\text{disease}\mid\text{+}) &= \frac{P(\text{+}\mid\text{disease}) P(\text{disease})}{P(+)} \\
 &= \frac{P(\text{+}\mid\text{disease}) P(\text{disease})}{P(\text{+}\mid\text{disease}) P(\text{disease}) + P(\text{+}\mid\text{healthy}) P(\text{healthy})} \\
 &= {\pi_1p}/[{\pi_1p} + {\pi_2{(1-p)}}]
\end{aligned}
$$
<br />

#### 3. 

* Injury

```{r echo=T}
table <- matrix(data = c(1601, 510, 162527, 412368), 2, 2, 
                dimnames = list(c("None", "Seat belt"), c("Fatal", "Nonfatal")));
pi_table <- prop.table(as.matrix(table), margin = 1)
difference_of_proportions <- pi_table[1,1] - pi_table[2,1];
relative_risk <- pi_table[1,1] / pi_table[2,1];
odds_ratio <- (pi_table[1,1] * pi_table[2,2]) / (pi_table[1,2] * pi_table[2,1]);
```
* difference of proportions = `r difference_of_proportions`
* relative risk = `r relative_risk`
* odds ratio = `r odds_ratio`
* Because both $\pi_{11}$ and ${\pi_{12}}$ are close to zero
<br />

#### 4a. 
```{r echo=T}
table <- matrix(data = c(28, 18, 656, 658), 2, 2, 
                dimnames = list(c("placebo", "Aspirin"), c("Yes", "No")));
pi_table <- prop.table(as.matrix(table), margin = 1)

sample_odds_ratio <- (table[1,1] * table[2,2]) / (table[2,1] * table[1,2]);

se_log_theta <- sqrt(sum(1/table));

CI_log_theta <- qnorm(c(.025, .975), log(sample_odds_ratio), se_log_theta);
CI_theta <- exp(CI_log_theta);
```

* Sample odds ratio is `r sample_odds_ratio`
* standard error of $log \hat{\theta}$ is `r se_log_theta`
* 95% confidence interval for $log \theta$ lies between `r CI_log_theta[1]` and `r CI_log_theta[2]`
* 95% confidence interval for $\theta$ lies between `r CI_theta[1]` and `r CI_theta[2]`
* The 95% confidence interval includes the value 1. Hence it is possible for the two to be equal.
<br />

#### 4b. 
```{r}
exp_diff <- pi_table[1,1] - pi_table[2,1];
var <- pi_table[1,1] * pi_table[1,2] / sum(table[1,]) +
  pi_table[2,1] * pi_table[2,2] / sum(table[2,]);
sd <- sqrt(var);
wald_CI <- qnorm(c(.025, .975), exp_diff, sd);
```
* 95% confidence interval for difference of proportions lies between `r wald_CI[1]` and `r wald_CI[2]`
<br />

#### 4c. 
````{r}
sample_relative <- pi_table[1,1] / pi_table[2,1];
exp_log_relative <- log(sample_relative);
var <- ((1 - pi_table[1,1]) / ((sum(table[1,])) * pi_table[1,1])) +
  ((1 - pi_table[2,1]) / ((sum(table[2,])) * pi_table[2,1]))
sd <- sqrt(var);
wald_CI_log <- qnorm(c(.025, .975), exp_log_relative, sd);
wald_CI <- exp(wald_CI_log);
```
* Sample relative risk = `r sample_relative`
* 95% confidence interval for the log relative risk lies between `r wald_CI_log[1]` and `r wald_CI_log[2]`
* 95% confidence interval for the relative risk lies between `r wald_CI[1]` and `r wald_CI[2]`
<br />


#### 5. 

$H_0$: independence for all combination
$H_1$: at least one combination not independent
```{r echo=T}
table <- matrix(data = c(178, 570, 138, 138, 648, 252, 108, 442, 252), 3, 3, 
                dimnames = list(c("Less than high school", "High school or junior college",
                                  "Bachelor or graduate"),
                                c("Fundamentalist", "Moderate", "Liberal")));

chisq.test(table)
```
* p-value is zero; there are enough evidence to reject $H_0$; there is an association.