---
title: <center><h1> 2018R2 Data Mining (STAT5104)  Assignment 1 Q2</h1></center><br />
author: <center>Yiu Chung WONG 1155017920</center>
output:
  pdf_document: default
  html_document:
    keep_md: yes
    code_folding:
  word_document: default
--- 

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
```

```{r echo = TRUE, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
set.seed(12345)
```

###a)
```{r read data}
dc <- read.csv("tele.csv", header = TRUE, sep = ",")            #read data
inTrain <- createDataPartition(dc$Change, p = .9) %>% unlist(.) #create index for train / test partition
d0 <- dc[inTrain,]                                              #select observations for training 
d1 <- dc[-inTrain,]                                             #select observations for testing
```

###b)
```{r ctree}
control <- rpart.control(maxdepth = 3)
ctree <- rpart(data = d0, formula = Change ~ ., control = control, method = 'class')
```

\newpage
###c)
```{r print}
rpart.plot(ctree)             #print plot
print(ctree)
asRules(ctree, compact=FALSE) #print rules

oneSum <- sum(d0$Change)      #calculate total number of observaation where Change == 1
                              #in the training set
zeroSum <- nrow(d0) - oneSum  #calculate total number of observaation where Change == 0
                              #in the training set
```

Rule Number 4: Day_Mins< 264.45 and CustServ_Calls< 3.5 then Change = 0<br />
Support = 2563 / 3288 = `r 2563 / 3288`<br />
Confidence = 1 - (198 / 2563) = `r 1 - (198 / 2563)`<br />
Capture = (2563 - 198) / 2542 = `r (2563 - 198)/zeroSum`<br />

Rule Number 10: Day_Mins< 264.45 and CustServ_Calls>= 3.5 and Day_Mins>=160.2 then Change = 0<br />
Support = 130 / 3288 = `r 130 / 3288`<br />
Confidence = 1 - (32 / 130) = `r 1 - (32 / 130)`<br />
Capture = (130 - 32) / 2542 = `r (130 - 32)/zeroSum`<br />

Rule Number 11: Day_Mins< 264.45 and CustServ_Calls>= 3.5 and Day_Mins< 160.2 then Change = 1<br />
Support = 88 / 3288 = `r 88 / 3288`<br />
Confidence = 1 - (10 / 88) = `r 1 - (10 / 88)`<br />
Capture = (88 - 10) / 418 = `r (130 - 32)/oneSum`

Rule Number 6: Day_Mins>=264.45 and Vmail_plan=yes then Change = 0<br />
Support = 44 / 3288 = `r 44 / 3288`<br />
Confidence = 1 - (6 / 44) = `r 1 - (6 / 44)`<br />
Capture = (44 - 6) / 2542 = `r (44 - 6)/zeroSum`

Rule Number 14: Day_Mins>=264.45 and Vmail_plan=yes and then Eve_Mins<187.75 Change = 0<br />
Support = 47 / 3288 = `r 47 / 3288`<br />
Confidence = 1 - (21 / 47) = `r 1 - (21 / 47)`<br />
Capture = (47 - 21) / 2542 = `r (47 - 21)/zeroSum`

Rule Number 15: Day_Mins>=264.45 and Vmail_plan=yes and then Eve_Mins>=187.75 Change = 1<br />
Support = 88 / 3288 = `r 88 / 3288`<br />
Confidence = 1 - (5 / 88) = `r 1 - (5 / 88)`<br />
Capture = (88 - 5) / 2542 = `r (88 - 5)/oneSum`

###d)
```{r}
test <- predict(ctree)                              #predict using training data 
cl_test <- max.col(test) - 1                        #rename columns from  1, 2 to 0, 1
test_table <- table(cl_test, d0$Change)             #put data into classification table 
test_table

validation <- predict(ctree, newdata = d1)          #predict using testing data 
cl_validation <- max.col(validation) - 1            #rename columns from  1, 2 to 0, 1
validation_table <- table(cl_validation, d1$Change) #put data into classification table 
validation_table
```
d0 error rate: (`r test_table[2]` + `r test_table[3]`) / `r sum(test_table)`  = `r (test_table[2] + test_table[3]) / sum(test_table)`

d1 error rate: (`r validation_table[2]` + `r validation_table[3]`) / `r sum(validation_table)`  = `r (validation_table[2] + validation_table[3]) / sum(validation_table)`